DATA 621 - Business Analytics and Data Mining

### 2020-12-13 Final_project/Final Project Paper - Group 2.Rmd
Clean up and reorder

<!-- The second is `exceedMean`. This will be a factor if the views for the video -->
<!-- are greater than the *mean* views for all of the extracted data set. This will -->
<!-- also indicate the skewness of the data. Were it symmetrical, and we know it is -->
<!-- not from the graphs above, the mean should be near the median and this factor -->
<!-- should be close to equal for both 0 and 1.  -->

@@

# Precaclulate the mean views
usMean <- mean(usExtract$views)

# Add the engineered feature
usExtract[, `:=`(balance = ifelse(rtgDisabled, 1,
                               pmin(likes, dislikes) / pmax(likes, dislikes)),
                 engagement = ifelse(rtgDisabled, 1, 
                                     comments / sum(likes, dislikes)))]
                 # exceedMean = factor(ifelse(views > usMean, 1, 0),
                 #                     levels = c(0, 1), labels = c(0, 1)))

```
```{r plotBalance, fig.height=4, fig.width=8}
ggplot(usExtract[, .(balance)], aes(x = balance)) +
  geom_density() +
  ggtitle("Distribution of Balance")

ggplot(usExtract[, .(engagement)], aes(x = engagement)) +
  geom_density() +
  scale_x_log10() +
  ggtitle("Distribution of Engagement Log Scale")

@@

It's pretty clear from the distribution of `balance` that, at least for videos
viewed in the US, there is a healthy dose of disagreement!

<!-- ```{r plotExceedMean, fig.height=4, fig.width=8} -->
<!-- # Melt for exceedMean -->
<!-- usSetF <- melt(usExtract, measure.vars = 'exceedMean', variable.factor = FALSE, -->
<!--                variable.name = 'measure', value.name = 'value') -->

<!-- # Plot the exceedMean variable -->
<!-- ggplot(usSetF[measure == 'exceedMean'], aes(x = value)) + -->
<!--   geom_bar() + -->
<!--   ggtitle("Proportion Above and Below Mean") + -->
<!--   scale_y_continuous(labels = comma) -->
<!-- ``` -->

<!-- As with many events which follow the Pareto rule, most of the views are  -->
<!-- attributable to a minority of the videos, as seen by the 0/1 disparity. -->

@@

What may be most interesting is that the disparity ratio indicator us a very
powerful and significant indicator. Conflict seems to be good for Youtube
videos. The more a video is argued over, the more views it seems to get!

@@

This approach *appears* similar to that of the common technique of performing a
standard linear regression on the logs of the observations, but is different. As
per [@Gelman2006Lta], one approach *"…log transforms observed values, while the*
*second one log transforms the expected value.…the key difference being the*
*relation between the predicted value and the variance."*

```{r lm2Train}
set.seed(181)
lm2 <- train(x = seenX, y = seenY, family = gaussian(link = 'log'),
             method = 'glmStepAIC', direction = 'both', trace = 0,
             trControl = trC)
```

```{r lm3Train_poi}
set.seed(9450)
lm3_test<- train(x = seenX, y = seenY, family = poisson(link='log'),
            method = 'glmStepAIC', direction = 'both', trace = 0,
            trControl = trC)
```

\footnotesize
```{r poi_coeff}
kable(summary(lm3_test$finalModel)$coefficients,
      caption = "Model 3 (Poisson) Linear Regression Output",
      digts = 3L, format.args = list(big.mark = ','))
```
\normalsize

\footnotesize
```{r lm2Table}
kable(summary(lm2$finalModel)$coefficients,
      caption = "Model 3 (Gaussian) Linear Regression Output",
      digts = 3L, format.args = list(big.mark = ','))
```
\normalsize

@@

Nevertheless, the *relative* magnitude and sign of the parameters are in line
with our expectations. Both `likes` and `dislikes` are positively correlated
with views and `comments` is negatively correlated. A few more categories join
`Film & Animation` as contributing to excess views over the baseline
`Entertainment`. But `Nonprofits & Activism` has the largest factor by far, and
it is negative, which stands in contrast to the second model. Lastly, having a
higher balance---ratio of likes to dislikes closer to 1---tends to increase
views. Conflict seems to be good for YouTubers!

@@

# Predict using the models
lm1P <- predict(lm1, newdata = hideX)
lm2P <- predict(lm2, newdata = hideX)
lm3P <- predict(lm3, newdata = hideX)
lm3_testP <- predict(lm3_test, newdata = hideX)
# Compare the results
compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'ElasticNet', 'GLM: Poisson+Log '),
                        RMSE = c(RMSE(lm1P, hideY), RMSE(lm2P, hideY),
                                 RMSE(lm3P, hideY), RMSE(lm3_testP, hideY)),
                        R2 = c(R2(lm1P, hideY, formula = 'traditional'),
                               R2(lm2P, hideY, formula = 'traditional'),
                               R2(lm3P, hideY, formula = 'traditional'),
                               R2(lm3_testP, hideY, formula = 'traditional')),
                        MAE = c(MAE(lm1P, hideY), MAE(lm2P, hideY),
                                MAE(lm3P, hideY), MAE(lm3_testP, hideY)))
kable(compTable, digits = 3L, format.args = list(big.mark = ','),
      caption = "Model Performance on Test Set")
```

@@

  * Make sure your videos will engender heated discussions. You want as many
  *dislikes* as likes. That seems to drive attention to your video!
  * Keep your ratings enabled. Allowing users to rate your video drives
  attention to them!
  * Perhaps disable comments, though.
  * Stay **away** from activism. 
  * Talk about films or comics. This probably includes the anime and manga
  phenomena.
  * Having more tags is better than having fewer.

@@

### Variable Importance: Model 2

```{r lm2VI}
# Model 2: VarImp 
ggplot(varImp(lm2)) + ggtitle('Variable Importance: Model 2')
```

### Variable Importance: Model 3
```{r lm3VI}
# Model 3: VarImp 
ggplot(varImp(lm3)) + ggtitle('Variable Importance: Model 3')
```

@@

<!-- Model Building -->
```{r holdout}
```
```{r dummyVarsAA}
```
```{r lm1Train}
```
```{r lm1Table}
```
```{r lm2Train}
```
```{r lm2Table}
```
```{r lm3Train}
```

@@

<!-- Statistical Appendix-->
```{r statsTable}
```
```{r lm1VI}
```
```{r lm2VI}
```
```{r lm2VI}
```
```{r lm3Plot}
```

>>>>

@@

# Add the engineered feature
usExtract[, `:=`(balance = ifelse(rtgDisabled, 1,
                               pmin(likes, dislikes) / pmax(likes, dislikes)),
                 engagement = ifelse(rtgDisabled, 1, 
                                     comments / (likes + dislikes)))]
engageMode <- prettyNum(density(usExtract$engagement, kernel = 'ep')$x[
  which.max(density(usExtract$engagement, kernel = 'ep')$y)], digits = 4L)
engageMean <- prettyNum(mean(usExtract$engagement), digits = 4L)
engageMax <- prettyNum(max(usExtract$engagement), digits = 4L)
```
```{r plotBalance, fig.height=4, fig.width=8}
ggplot(usExtract[, .(balance)], aes(x = balance)) +
  geom_density() +
  ggtitle("Distribution of Balance")

ggplot(usExtract[, .(engagement)], aes(x = engagement)) +
  geom_density() +
  ggtitle("Distribution of Engagement")

@@

It's pretty clear from the distribution of `balance` that, at least for videos
viewed in the US, there is a healthy dose of disagreement, as its mode is much
closer to 0 than 1. Similarly, `engagement` exhibits right-skewed behavior, with
a mean of `r engageMean` and a mode at `r engageMode`, yet a maximum of
`r engageMax`.

@@

What may be most interesting is that the disparity and engagement ratios are
powerful and significant indicators. Conflict seems to be good for Youtube
videos. The more a video is argued over, the more views it seems to get!

@@

The first approach *appears* similar to that of the common technique of
performing a standard linear regression on the logs of the observations, but it
is different. As per [@Gelman2006Lta], one approach *"…log transforms observed*
*values, while the second one log transforms the expected value.…the key*
*difference being the relation between the predicted value and the variance."*

The second approach is the classic Poisson regression.

```{r lm2aTrain}
set.seed(181)
lm2a <- train(x = seenX, y = seenY, family = gaussian(link = 'log'),
             method = 'glmStepAIC', direction = 'both', trace = 0,
             trControl = trC)
```

```{r lm2b}
set.seed(181)
lm2b<- train(x = seenX, y = seenY, family = poisson(link='log'),
            method = 'glmStepAIC', direction = 'both', trace = 0,
            trControl = trC)
```

\footnotesize
```{r lm2Tables}
kable(summary(lm2a$finalModel)$coefficients,
      caption = "Model 3a (Gaussian) Linear Regression Output",
      digts = 3L, format.args = list(big.mark = ','))
kable(summary(lm2b$finalModel)$coefficients,
      caption = "Model 3b (Poisson) Linear Regression Output",
      digts = 3L, format.args = list(big.mark = ','))
```
\normalsize

@@

Nevertheless, the *relative* magnitude and sign of the parameters are in line
with our expectations. Both `likes` and `dislikes` are positively correlated
with views and `comments` is negatively correlated. A few more categories join
`Film & Animation` as contributing to excess views over the baseline
`Entertainment`. But `Nonprofits & Activism` has the largest factor by far, and
it is negative, which stands in contrast to the second model. Lastly, having a
higher `balance`---ratio of likes to dislikes closer to 1--- or a higher
`engagement` tends to increase views. Conflict seems to be good for YouTubers!

@@

# Predict using the models
lm1P <- predict(lm1, newdata = hideX)
lm2aP <- predict(lm2a, newdata = hideX)
lm2bP <- predict(lm2b, newdata = hideX)
lm3P <- predict(lm3, newdata = hideX)
# Compare the results
compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'GLM: Poisson+Log',
                                  'ElasticNet'),
                        RMSE = c(RMSE(lm1P, hideY), RMSE(lm2aP, hideY),
                                 RMSE(lm2bP, hideY), RMSE(lm3P, hideY)),
                        R2 = c(R2(lm1P, hideY, formula = 'traditional'),
                               R2(lm2aP, hideY, formula = 'traditional'),
                               R2(lm2bP, hideY, formula = 'traditional'),
                               R2(lm3P, hideY, formula = 'traditional')),
                        MAE = c(MAE(lm1P, hideY), MAE(lm2aP, hideY),
                                MAE(lm2bP, hideY), MAE(lm3P, hideY)))
kable(compTable, digits = 3L, format.args = list(big.mark = ','),
      caption = "Model Performance on Test Set")
```

@@

  * Make sure your videos will engender heated discussions. You want as many
  *dislikes* as likes. That seems to drive attention to your video!
  * Keep your ratings enabled. Allowing users to rate your video drives
  attention to them!
  * Perhaps disable comments, though.
  * Stay **away** from activism. 
  * Talk about films or comics. This probably includes the anime and manga
  phenomena.
  * Talking about cars is also good.
  * Having more tags is better than having fewer.

@@

### Variable Importance: Model 2a

```{r lm2aVI}
# Model 2: VarImp 
ggplot(varImp(lm2a)) + ggtitle('Variable Importance: Model 2a')
```

### Variable Importance: Model 2b

```{r lm2bVI}
# Model 2: VarImp 
ggplot(varImp(lm2b)) + ggtitle('Variable Importance: Model 2b')
```

### Variable Importance: Model 3

```{r lm3VI}
# Model 3: VarImp 
ggplot(varImp(lm3)) + ggtitle('Variable Importance: Model 3')
```

@@

<!-- Model Building -->
```{r holdout}
```
```{r dummyVarsAA}
```
```{r lm1Train}
```
```{r lm1Table}
```
```{r lm2aTrain}
```
```{r lm2bTrain}
```

```{r lm2Tables}
```
```{r lm3Train}
```

@@

<!-- Statistical Appendix-->
```{r statsTable}
```
```{r lm1VI}
```
```{r lm2aVI}
```
```{r lm2bVI}
```
```{r lm3VI}
```
```{r lm3Plot}
```
